---
title: "sept17_data_wrangling_i"
output: github_document
---

```{r}
library(tidyverse)
```

load in a dataset: 

relative path: starts from where you are and then goes to file location 
eg: "./data/data_import_examples/FAS_litters.csv" when we are in data_wrangling_i 


```{r}
## reads in a dataset 
litters_data = read_csv(file = "./data/data_import_examples/FAS_litters.csv")

litters_data 

```

print data = good 1st step (onlt first few rows and columns)

to take janitor clean names and show them in the file 
```{r}
litters_data = janitor::clean_names(litters_data)

litters_data
```

loading pups_data: 

```{r}
## reads in a dataset 
pups_data = read_csv(file = "./data/data_import_examples/FAS_pups.csv")

pups_data 

#clean name function is in the janitor package 

pups_data = janitor::clean_names(pups_data)

pups_data

```

only use view function on console - not rmd file 

```{r}
head(pups_data)

tail(pups_data)

library(skimr)

skim(pups_data)

#skim function is in the skimr package and provides skim stats about the data

```


Parsed with column specification:
cols(
  Group = col_character(),
  `Litter Number` = col_character(),
  `GD0 weight` = col_double(),
  `GD18 weight` = col_double(),
  `GD of Birth` = col_double(),
  `Pups born alive` = col_double(),
  `Pups dead @ birth` = col_double(),
  `Pups survive` = col_double()
)

^tells the type of data in each column 


You can use the readeacel package to load excel datsets as well 

```{r}
library(readxl)

mlb11_data = read_excel(path = "./data/data_import_examples/mlb11.xlsx")

mlb11_data

```

Read in SAS 

```{r}

pulse_data = haven::read_sas("./data/data_import_examples/public_pulse_data.sas7bdat")

pulse_data


```


know the difference between read_csv and read.csv 

factor variable vs character var 


exporting data: 
```{r}
library(readxl)

mlb11_data_subet = read_excel(
  path = "./data/data_import_examples/mlb11.xlsx", 
                        range = "A1:D7")

write( )

mlb11_data

```








